---
author: "Mara Childs"
content: "Top Tracks 2017 Analysis"
date: "3/18/2018"
output:
  html_document:
    css: style.css
---

<!-- Using "include = FALSE" to hide code work -->
```{r libraries, include = FALSE}
library(httr)
library(dplyr)
library(knitr)
library(plotly)
```

# Spotify's Top 100 Tracks of 2017
##### An in-depth look at what we streamed the most

For this project, I'm exploring the data from Spotify's top streamed tracks of 2017. [This Kaggle dataset](https://www.kaggle.com/nadintamer/top-tracks-of-2017/data) includes Spotify's track metadata as well as 13 different track metrics for a total of 16 data points per track: track id, track name, track artist, danceability, energy, key, loudness, mode, speechiness, acousticness, instrumentalness, liveness, valence, tempo, duration (ms), and time signature. Spotify provides a brief interpretation of these descriptors on its [developer API endpoint reference](https://developer.spotify.com/web-api/get-audio-features/).

```{r data, include = FALSE}
# Read in Spotify Top Songs 2017 csv
tracks <- read.csv("data/toptracks2017.csv", stringsAsFactors = FALSE)
```

```{r artists, include = FALSE}
# ARTISTS
artists <- tracks %>% select(name, artists)

# Tracks per artist
artist_popularity <- artists %>% select(artists) %>% count(artists) %>% arrange(-n)
top_artists <- artist_popularity %>% filter(n > 1)
```

### Artists
The top 100 tracks may be 100 individual songs, but there were `r nrow(topArtists)` artists with more than one track on the list. Topping the list with `r topArtists[1,2]` tracks each were `r topArtists[1,1]` and `r topArtists[2,1]`. 

```{r artist table, echo = FALSE}
# Top artists table
artist_col <- c("Artist Name","Number of Tracks")
kable(top_artists, row.names = NA, col.names = artist_col, caption = "Artists With More Than One Track")
```

